UKB Models

### 2024-05-17 eda/ukb_curation-eda.Rmd
cleaning up

```{r}
cn_data <- fread(file= paste0(base_path, "data/lifespan_dataset_dkatlas_CNonly.csv"))
#table(cn_data$study)
names(cn_data)
```

@@

```{r}
ukb_df <- cn_data %>%
  dplyr::filter(study == "UKB") %>%
  mutate(age_yrs = round(age_days/365.25),
         log_age = log(age_days, base=10),
         TBV=(sGMV+GMV+WMV+Ventricles),
         fs_version = as.factor(fs_version),
         sex = as.factor(sex),
         sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)),
         sex.age = sexMale * age_days) %>%
  na.omit() #1 subj has 1 missing val

@@

Add brainwide phenotypes!
Summing across regions for volumes (rather than summing GMV, sGMV, etc) for consistency with SA and CT

```{r}
vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]

ukb_df_totals <- ukb_df %>%
  mutate(Vol_total=rowSums(dplyr::select(., .dots=all_of(c(vol_list_regions)))),
         SA_total=rowSums(dplyr::select(., .dots=all_of(c(sa_list)))),
         CT_total=rowSums(dplyr::select(., .dots=all_of(c(ct_list)))))

#write data
write_csv(ukb_df_totals, file=paste0(base_path,"data/ukb_CN_data.csv"))
```

@@

```{r}
ukb_df_totals %>%
  dplyr::filter(age_yrs <= 49) %>%
ggplot(data=, aes(x=age_days)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)

ukb_df_totals %>%
  dplyr::filter(age_yrs >= 79) %>%
ggplot(data=, aes(x=age_days)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)
```

@@

```{r}
ukb_df_age.filt <- ukb_df_totals %>%
  dplyr::filter(age_yrs >= 50 & age_yrs <= 80)
#write data
fwrite(ukb_df_age.filt, file=paste0(base_path,"data/ukb_CN_data_agefilt.csv"))

nrow(ukb_df_totals) - nrow(ukb_df_age.filt)
```

@@

```{r}
ggplot(data=ukb_df_totals, aes(x=age_days, y=Vol_total)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity")

ggplot(data=ukb_df_totals, aes(x=age_days, y=TBV)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity") #TBV includes ventricles, WMV, subcortical GMV

ggplot(data=ukb_df_totals, aes(x=age_days, y=CT_total)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity")

ggplot(data=ukb_df_totals, aes(x=age_days, y=SA_total)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity")
```

@@

center
```{r}
ukb_centered <- ukb_df_totals %>%
  group_by(sex) %>%
  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
  ungroup()
```

@@

#write out
write_csv(ukb_df_sites, file=paste0(base_path, "data/ukb_to_model/ukb_CN_data_simsites.csv"))
```

```{r}
table(ukb_df_sites$sim.site)
table(ukb_df_sites$sim.site, ukb_df_sites$sex)
sum(is.na(ukb_df_sites$sim.site)) #all subj assigned to sites
anova(lm(age_days ~ sim.site, ukb_df_sites))
```

Need to come up with cleaner way to do this...

```{r}
site_sum_df <- ukb_df_sites %>%
  group_by(sim.site) %>%
  summarise(proportion_male = mean(sex == "Male"),
            min_age = min(age_days), max_age = max(age_days), mean_age = mean(age_days), 
            var_gmv = var(GMV), var_wmv=var(WMV), var_sgmv=var(sGMV), var_ven = var(Ventricles),
            sd_below.gmv = mean(GMV) - sd(GMV), sd_above.gmv = mean(GMV) + sd(GMV))
```

Center and visualize distributions

@@

##Exploratory QC

Adding euler number (from jakob) back into combat-ed dfs so i can test how controlling for euler in centile scores changes the performance of combat configs (esp. in CT based on Bedford et al preprint).

```{r}
cn.df_euler <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly_euler.csv"), stringsAsFactors = TRUE)

ukb.df_euler <- cn.df_euler %>%
  dplyr::filter(study == "UKB") %>%
  dplyr::select(participant, surfholes)

table(unique(ukb_df_age.filt$participant) %in% unique(ukb.df_euler$participant)) #all subjects in df!

fwrite(ukb.df_euler, file=paste0(base_path,"data/ukb_euler_nums.csv"))
```

>>>>

```{r}
cn_data <- fread(file= paste0(base_path, "data/lifespan_dataset_dkatlas_CNonly.csv"))
#table(cn_data$study)
#names(cn_data)
```

@@

```{r}
ukb_df <- cn_data %>%
  dplyr::filter(study == "UKB") %>%
  mutate(age_yrs = round(age_days/365.25),
         log_age = log(age_days, base=10),
         fs_version = as.factor(fs_version),
         sex = as.factor(sex),
         sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)),
         sex.age = sexMale * age_days) %>%
  na.omit() #1 subj has 1 missing val

@@

@@

```{r}
ukb_df %>%
  dplyr::filter(age_yrs <= 49) %>%
ggplot(data=, aes(x=age_days)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)

ukb_df %>%
  dplyr::filter(age_yrs >= 79) %>%
ggplot(data=, aes(x=age_days)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)
```

@@

```{r}
ukb_df_age.filt <- ukb_df %>%
  dplyr::filter(age_yrs >= 50 & age_yrs <= 80)
#write data
fwrite(ukb_df_age.filt, file=paste0(base_path,"data/ukb_CN_data_agefilt.csv"))

nrow(ukb_df) - nrow(ukb_df_age.filt)
```

@@

@@

center
```{r}
ukb_centered <- ukb_df %>%
  group_by(sex) %>%
  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
  ungroup()
```

@@

#write out
fwrite(ukb_df_sites, file=paste0(base_path, "data/ukb_to_model/ukb_CN_data_simsites.csv"))
```

```{r}
table(ukb_df_sites$sim.site)
table(ukb_df_sites$sim.site, ukb_df_sites$sex)
sum(is.na(ukb_df_sites$sim.site)) #all subj assigned to sites
anova(lm(age_days ~ sim.site, ukb_df_sites))
```
Center and visualize distributions

@@
